Reconstructing Propositional Proofs in Type Theory

### 2017-10-26 paper/sections/metis-language-and-proofs.tex
[ paper ] lots of changes.

\Metis is an automatic theorem prover written in \len{ML} for
first-order logic with equality developed by
\citeauthor{hurd2003first}~\cite{hurd2003first}. This prover is
suitable for proof reconstruction since it provides well-documented
proofs to justify its deduction steps from the basis of only six
inference rules. In Fig.~\ref{fig:metis-inferences}, we show the three
\Metis rules for propositional logic.

@@

\caption{\Metis inference rules.}
  \begin{center}
  {\renewcommand{\arraystretch}{1.6}%
    \label{tab:agda-metis-table}
    \begin{tabular}{|@{\hspace{2mm}}l@{\hspace{2mm}}l@{\hspace{2mm}}c@{\hspace{2mm}}|}
    \hline
    \textbf{\Metis rule name} & \textbf{Purpose} &\textbf{Theorem}\\ \hline

>>>>

\Metis is an automatic theorem prover for first-order logic with equality developed by
 \citeauthor{hurd2003first}~\cite{hurd2003first}.
This prover is suitable for proof reconstruction since it provides well-documented proofs to justify its deduction steps from the basis of only six inference rules. In Fig.~\ref{fig:metis-inferences}, we show the three \Metis rules for propositional logic.

@@

\caption{\Metis inference rules.}
  \begin{center}
  {\renewcommand{\arraystretch}{1.6}%
    \label{tab:agda-metis-table}
    \begin{tabular}{|@{\hspace{2mm}}l@{\hspace{2mm}}l@{\hspace{2mm}}c@{\hspace{2mm}}|}
    \hline
    \textbf{\Metis Rule} & \textbf{Purpose} &\textbf{Theorem}\\ \hline

### 2017-10-26 paper/sections/proof-reconstruction.tex
[ paper ] lots of changes.

The proof-reconstruction approach proposed consists of a series
of steps similar to the workflow presented by \citeauthor{sultana2015} in
\cite{sultana2015}. This process is a translation from a source
system to a target system. In our case, the system of origin, the
automatic theorem prover, is \Metis; the target system is a proof
assistant, \Agda. We choose \Agda, but another proof assistant with the
same support of type theory and inductive types could be used.

@@

Now, we have finished the formalization of every inference rule in
a \Metis proof, we are able to justify step-by-step any proof
for a problem in propositional logic.
For instance, we tested successfully the translation by \Athena jointly with
the \Agda formalizations of the rules mentioned above with more than eighty
representative theorems in propositional logic. An interested reader can test
the problems~\cite{Prieto-Cubides2017} in the \Athena tool
repository~\cite{Athena}.

>>>>

The proof-reconstruction approach proposed consists of a series
of steps similar to the workflow presented by
\citeauthor{sultana2015}~\cite{sultana2015}.
This process is a translation from a source
system to a target system. In our case, the system of origin, the
automatic theorem prover, is \Metis; the target system is a proof
assistant, \Agda. We choose \Agda, but another proof assistant with the
same support of type theory and inductive types could be used.

@@

We have finished the formalization of every inference rule in
a \Metis proof, we are able to justify step-by-step any proof
for a problem in propositional logic.
For instance, we tested successfully the translation by \Athena jointly with
the \Agda formalizations of the rules mentioned above with more than eighty
representative theorems in propositional logic. An interested reader can test
the problems~\cite{Prieto-Cubides2017} in the \Athena tool
repository~\cite{Athena}.

### 2017-10-26 paper/sections/proof-reconstruction/clausify.tex
[ paper ] added some corrections.
[ paper ] lots of changes.

The \clausify rule transforms a
propositional formula into its clausal normal form, a conjunction
of clauses. A \emph{clause} is the disjunction of zero or more
literals.
It is important to notice that this kind of conversion between one formula
to its clausal normal form is not unique, and \Metis has a customized
approach to perform that transformation. That is the reason we perform some
reordering for the \CNF of the source formula to fulfill this issue.

\begin{myexample}
In the following \Metis \TSTP derivation, we see how
\clausify expands the\texttt{norm0} formula to get \texttt{norm1} formula
by using distributive laws included in the conjunctive normal form conversion.

\begin{verbatim}
fof(a1, axiom, p ⇒ (q ∧ r)).
...
fof(norm0, ¬ p ∨ (q ∧ r), inf(canonicalize, [a1])).
fof(norm1, (¬ p ∨ q) ∧ (¬ p ∨ r), inf(clausify, [norm0])).
...
\end{verbatim}

@@

\begin{myremark}
The \clausify rule is often preceded by \canonicalize.
Both rules perform \emph{Clausification} that introduces
into the problem axioms and definitions in the domain of the solution.
The clausification algorithms is mainly described on paper by
\citeauthor{Sutcliffe1996}~\cite{Sutcliffe1996}.
\end{myremark}

>>>>

The \clausify rule transforms a
propositional formula into its clausal normal form.
It is important to notice that this kind of conversion between one formula
to its clausal normal form is not unique, and \Metis has a customized
approach to perform that transformation. That is the reason we perform some
reordering for the \CNF of the source formula to fulfill this issue.

\begin{myexample}
In the following \TSTP derivation by \Metis, we see how
\clausify transform the \texttt{norm0} formula to get \texttt{norm1} formula.

\begin{verbatim}
fof(norm0, ¬ p ∨ (q ∧ r) ...
fof(norm1, (¬ p ∨ q) ∧ (¬ p ∨ r), inf(clausify, [norm0])).
\end{verbatim}

@@

\begin{myremark}
The \clausify rule is often preceded by the \canonicalize rule.
Both rules perform \emph{Clausification} that introduces
into the problem axioms and definitions in the domain of the solution.
The clausification is widely described by
\citeauthor{Sutcliffe1996}~\cite{Sutcliffe1996}.
\end{myremark}
